```{r}
source("requirements.R")
source("methods.R")
source("sim.R")
```

Define a function to extract place field properties from the output of `sim`.

```{r}
get_mean_place_field_properties = function(place_cell_list) {
    centroids_day0 = get_centroids(place_cell_list[[1]]$firing_rates)
    drift = 
      sapply(1:length(place_cell_list), function(i) {
        centroids_day_i = get_centroids(place_cell_list[[i]]$firing_rates)
        mean(abs(centroids_day_i - centroids_day0), na.rm=T)
      })
    
    median_drift = 
      sapply(1:length(place_cell_list), function(i) {
        centroids_day_i = get_centroids(place_cell_list[[i]]$firing_rates)
        median(abs(centroids_day_i - centroids_day0), na.rm=T)
      })
    
    total_place_cells = 
      sapply(1:length(place_cell_list), function(i) {
        get_centroids(place_cell_list[[i]]$firing_rates) %>% 
          na.omit() %>% 
          length()
      })
    
    recurring_place_cells = 
      sapply(1:length(place_cell_list), function(i) {
        get_centroids(place_cell_list[[i]]$firing_rates) %>% 
          {.[!is.na(centroids_day0)]} %>%
          na.omit() %>% 
          length()
      })
    
    data.frame(
      Drift = drift,
      Median_drift = median_drift,
      Total_PCs = total_place_cells,
      Recurring_PCs = recurring_place_cells,
      Day = 0:(length(place_cell_list)-1)
    )
}
```

Run simulations at a wide range of learning rates to evaluate the effects of learning rate on place field properties.

```{r}
set.seed(3743113)

eta = c(1, 1e-1, 1e-2, 1e-3, 1e-4, 1e-5, 1e-6, 1e-7, 0) %>% replicate(10, .) %>% c()
df_by_learning_rate = 
  sapply(eta, function(eta) {
    sim(learning_rate=eta, dropout_fun="E%-max") %>%
      get_mean_place_field_properties() %>%
      mutate(Eta = eta)
  }, simplify=F) %>%
  bind_rows()

# save the result because this simulation is slow
write.csv(df_by_learning_rate, "../Simulation_data/place_cells_by_learning_rate_052818.csv", row.names=F)
```

Generate figures showing how place field properties change with the learning rate.

```{r}
# load the saved result of the above simulation
df_by_learning_rate = read.csv("../Simulation_data/place_cells_by_learning_rate_052818.csv")

# plot median drift
g1 = 
  df_by_learning_rate %>%
  mutate(Eta = factor(Eta)) %>%
  ggplot(aes(Day, Median_drift)) +
  facet_grid(~Eta) +
  geom_line(aes(group=1), stat="summary", fun.data="mean_se") +
  geom_errorbar(stat="summary", fun.data="mean_se") +
  geom_point(size=.1, pch=16, alpha=.5) +
  annotate("point", x=17.5, y=2.5, shape=13, size=3, color="red") +
  ylab("Place field drift [cm]") +
  ggtitle("Learning rate") +
  theme(plot.title = element_text(hjust=.5, vjust=-1, size=8))

# plot the total number of place cells
g2 = 
  df_by_learning_rate %>%
  mutate(Eta = factor(Eta)) %>%
  ggplot(aes(Day, Total_PCs)) +
  geom_abline(slope=0, intercept=400, lty=2, color="red") +
  geom_abline(slope=0, intercept=200, lty=3, color="red") +
  facet_grid(~Eta) +
  geom_line(aes(group=1), stat="summary", fun.data="mean_se") +
  geom_errorbar(stat="summary", fun.data="mean_se") +
  geom_point(size=.1, pch=16, alpha=.5) +
  ylab("Total PCs [count]") +
  ggtitle("Learning rate") +
  theme(plot.title = element_text(hjust=.5, vjust=-1, size=8))

# plot the number of recurring place cells
g3 =
  df_by_learning_rate %>%
  mutate(Eta = factor(Eta)) %>%
  group_by(Eta, replicate_n) %>%
  ddply(c("Eta", "replicate_n"), function(df) {
    df$Original = df$Total_PCs[df$Day==0]
    df
  }) %>%
  mutate(Recurring_fraction = Recurring_PCs/Original) %>%
  ggplot(aes(Day, Recurring_fraction)) +
  annotate(
    "line", color="red", lty=2,
    x=c(0, 5, 10, 15, 20, 25, 30),
    y=c(1, .25, .20, .18, .15, .14, .13)
  ) +
  facet_grid(~Eta) +
  geom_line(stat="summary", fun.data="mean_se") +
  geom_errorbar(stat="summary", fun.data="mean_se") +
  geom_point(size=.1, pch=16, alpha=.5) +
  labs(y="Recurrence probability", title="Learning rate") +
  theme(plot.title = element_text(hjust=.5, vjust=-1, size=8))
```

Run a seperate simulation to find the average synaptic input per place cell at various learning rates.

```{r}
set.seed(431432)

X = load_input() # grid cells

# run simulation to find average input
df_mean_input = 
  sapply(c(0, 1e-7, 1e-6, 1e-5, 1e-4, 1e-3, 1e-2, 1e-1, 1), function(eta) {
    mean_input =
      sim(learning_rate=eta, dropout_fun="E%-max") %>%
      {np$dot(X, .[[61]]$weights)} %>% 
      colMeans()
    
    df = 
      data.frame(
        Mean_input = mean_input,
        Eta = eta
      )
    
    df
  }, simplify=F) %>%
    bind_rows()

# plot average synaptic input
g4 =
  df_mean_input %>%
  ggplot(aes(Mean_input)) +
  facet_grid(~Eta) +
  geom_histogram(bins=30, fill="black", color=NA) +
  labs(x="Synaptic input [au]", y="Count", title="Learning rate") +
  theme(plot.title = element_text(hjust=.5, vjust=-1, size=8))
```

Generate a figure combining all results.

```{r}
png("Model2_by_eta.png", width=6.18, height=5.5, units="in", res=300)
plot_grid(
  g1, g2, g3, g4, nrow=4,
  labels = c("A", "B", "C", "D"), label_size = 10
  )
dev.off()
```




