```{r}
source("requirements.R")
source("methods.R")
```

Create a function to simulate place cell activity in multiple environments defined by their unique grid cell activities.

```{r}
sim_multiple_environments = function(learning_rate, dropout_fun="E%-max", n_environments=3) {
  # values to save
  place_cell_list = lapply(1:n_environments, function(i) list())
  # parameters
  X = # grid cell firing rates
    lapply(1:n_environments, function(i) {
      load_input()[, sample(1:10000)]
    })
  
  n_units = 2000 # number of place cells
  n_inputs = 10000 # number of grid cells
  # initialize place cells
  place_cells = initialize_place_cells(n_units, n_inputs)
  # day zero
  for (env in 1:n_environments) {
    place_cells = calculate_firing_rates(place_cells=place_cells, X=X[[env]], dropout_fun=dropout_fun) # early phase
    place_cells = update_place_cell_weights(place_cells, X=X[[env]], learning_rate=learning_rate)
    place_cells = calculate_firing_rates(place_cells=place_cells, X=X[[env]], dropout_fun=dropout_fun) # late phase
    place_cell_list[[env]][[1]] = place_cells # save place cell data
  }
  # days 1 to 60
  for (day in 1:60) {
    place_cells = turnover(place_cells) # synapse turnover
    for (env in 1:n_environments) {
      place_cells = calculate_firing_rates(place_cells=place_cells, X=X[[env]], dropout_fun=dropout_fun) # early phase
      place_cells = update_place_cell_weights(place_cells, X=X[[env]], learning_rate=learning_rate)
      place_cells = calculate_firing_rates(place_cells=place_cells, X=X[[env]], dropout_fun=dropout_fun) # late phase
      place_cell_list[[env]][[day+1]] = place_cells # save place cell data
    }
  }
  return(place_cell_list)
}
```

Create a function to extract place field properties.

```{r}
get_mean_place_field_properties_multiple_environments = function(place_cell_list_all) {
  sapply(1:length(place_cell_list_all), function(i) {
    place_cell_list = place_cell_list_all[[i]]
    
    centroids_day0 = get_centroids(place_cell_list[[1]]$firing_rates)
    drift = 
      sapply(1:length(place_cell_list), function(i) {
        centroids_day_i = get_centroids(place_cell_list[[i]]$firing_rates)
        mean(abs(centroids_day_i - centroids_day0), na.rm=T)
      })
    
    median_drift = 
      sapply(1:length(place_cell_list), function(i) {
        centroids_day_i = get_centroids(place_cell_list[[i]]$firing_rates)
        median(abs(centroids_day_i - centroids_day0), na.rm=T)
      })
    
    total_place_cells = 
      sapply(1:length(place_cell_list), function(i) {
        get_centroids(place_cell_list[[i]]$firing_rates) %>% 
          na.omit() %>% 
          length()
      })
    
    recurring_place_cells = 
      sapply(1:length(place_cell_list), function(i) {
        get_centroids(place_cell_list[[i]]$firing_rates) %>% 
          {.[!is.na(centroids_day0)]} %>%
          na.omit() %>% 
          length()
      })
    
    data.frame(
      Drift = drift,
      Median_drift = median_drift,
      Total_PCs = total_place_cells,
      Recurring_PCs = recurring_place_cells,
      Day = 0:(length(place_cell_list)-1),
      Environment = i
    )
  }, simplify=F) %>%
    bind_rows()
}
```

Extract place cell properties from the multi-enviroment simulation.

```{r}
eta = c(0, 1e-4) %>% replicate(10, .) %>% c()
df_multiple_envs = 
  sapply(eta, function(eta) {
    place_cell_list = sim_multiple_environments(learning_rate=eta, dropout_fun="E%-max", n_environments=3)
    df_multiple_envs = get_mean_place_field_properties_multiple_environments(place_cell_list)
    df_multiple_envs %<>% mutate(Learning_rate = eta)
    df_multiple_envs
  }, simplify=F) %>%
  bind_rows()

# save results
write.csv(df_multiple_envs, "../Simulation_data/place_cells_in_multiple_env.csv", row.names=F)
```

Plot place cell offset over time.

```{r, fig.height=3, fig.width=2}
g_1 =
  df_multiple_envs %>%
  mutate(Environment = Environment %>% {
    paste("Env.", .)
  }) %>%
  mutate(Condition = ifelse(Learning_rate==0, "No learning", "Learning")) %>%
  mutate(Condition = factor(Condition, levels=c("No learning", "Learning"))) %>%
  ggplot(aes(Day, Median_drift, color=Condition, group=Condition)) +
  facet_grid(Environment~.) +
  geom_line(stat="summary", fun.data="mean_se") +
  geom_errorbar(stat="summary", fun.data="mean_se") +
  geom_point(size=.2, alpha=.1) +
  scale_color_aaas() +
  theme(strip.background = element_blank()) +
  ylab("Place field drift [cm]") +
  ylim(0, NA) +
  theme(
    legend.position = c(.7, .83),
    legend.margin = margin(0,0,0,0),
    legend.title = element_blank()
    ) +
  guides(color=guide_legend(keywidth = .5, keyheight = .5, nrow=2)) +
  annotate("point", x=17.5, y=2.5, shape=13, size=3)
```

Generate a representative multi-environment simulation.

```{r}
rep_learning_3_env = sim_multiple_environments(learning_rate=1e-4, dropout_fun="E%-max", n_environments=3)
```

Generate a raster plot of the representative simulation.

```{r}
df_rep_learning_3_env =
  sapply(1:3, function(env) {
    centroids_day0 = get_centroids(rep_learning_3_env[[env]][[1]]$firing_rates)
    idx = order(centroids_day0, na.last = NA)
    
    sapply(1:3, function(env_2) {
      sapply(1:61, function(day) {
        rep_learning_3_env[[env_2]][[day]]$firing_rates[, idx] %>%
          melt() %>%
          mutate(Day = day-1)
      }, simplify = F) %>%
        bind_rows() %>%
        mutate(Environment = env_2)
    }, simplify = F) %>%
        bind_rows() %>%
        mutate(Sorted_by = env)
    
  }, simplify = F) %>%
    bind_rows()

g_rep_3_env =
  df_rep_learning_3_env %>%
  filter(value != 0) %>%
  filter(Day %in% c(0, 5, 10, 15, 20, 25, 30, 60)) %>%
  mutate(Day = paste("Day", Day)) %>%
  mutate(Day = factor(Day, levels=c(
    paste("Day", c(0, 5, 10, 15, 20, 25, 30, 60))
  ))) %>%
  mutate(Sorted_by = paste("Sorted by\nenv.", Sorted_by)) %>%
  mutate(Sorted_by = factor(Sorted_by, levels=c(
    paste("Sorted by\nenv.", c(1, 2, 3))
  ))) %>%
  ggplot(aes(Var1, Var2, fill=factor(Environment))) +
  facet_grid(Sorted_by~Day, scales="free_y") +
  geom_raster(interpolate=T) +
  scale_fill_manual(values=c("grey20", "purple", "cyan3")) +
  scale_x_continuous(breaks=c(25, 75)) +
  theme(
    strip.background = element_blank(),
    legend.position = "none"
    ) +
  ylab("Place cell index") +
  xlab("Position [cm]")
```

Generate a figure combining all results.

```{r}
theme_set(
  theme_classic() +
    theme(
      text=element_text(size=8, family="Times New Roman"),
      legend.background=element_rect(fill=NA),
      legend.margin = margin(0, 0, 0, 0),
      axis.line = element_line(size=.25),
      axis.ticks = element_blank(),
      axis.text.y = element_text(margin = margin(r = -1)),
      axis.text.x = element_text(margin = margin(t = -1)),
      strip.background = element_rect(fill=NA, color=NA)
      )
)

png("multiple_envs.png", width=6.18, height=2.5, units="in", res=300)
plot_grid(g_1, g_rep_3_env, 
          rel_widths = c(1, 2),
          labels=c("A", "B"), label_size = 10
          )
dev.off()
```



