
This file is meant to be run after `Orientation Tuning Notebook.Rmd` and relies on variable and dependencies described and instantiated there.

Show firing rates by orientation for simulated V1 neurons. The resulting dataframe will be used to generate final figures. 

*Note that day 60 in this file is later interpretted as day 118 in other R script files and in the manuscript.*

```{r, fig.height=.5, fig.width=2}
y_original = 
  np$dot(y_LGN, w_LGN_to_V1_original) %>% 
  apply(1, function(x) ifelse(x>.9*max(x), x, 0)) %>% t()
y_final = 
  np$dot(y_LGN, w_LGN_to_V1) %>% 
  apply(1, function(x) ifelse(x>.9*max(x), x, 0)) %>% t()

df_original = 
  y_original %>%
  as.data.frame() %>%
  mutate(Theta = all_theta) %>%
  gather(Cell, Rate, -Theta) %>%
  mutate(Day = "0")

df_final = 
  y_final %>%
  as.data.frame() %>%
  mutate(Theta = all_theta) %>%
  gather(Cell, Rate, -Theta) %>%
  mutate(Day = "60")

df = bind_rows(df_original, df_final)

# filter out neurons that weren't active on the last day
df %<>% 
  group_by(Cell) %>%
  mutate(Remove = sum(Rate[Day=="60"]) == 0) %>%
  filter(Remove == F) %>%
  ungroup()

# show some example neurons
df %>%
  filter(Cell %in% sample(unique(Cell), 5)) %>%
  mutate(Degrees = (Theta + pi) * (360/(2*pi)) - 90) %>%
  ggplot(aes(Degrees, Rate, group=Day, color=Day)) +
  facet_grid(.~Cell) +
  geom_line(stat="summary", fun.data="mean_se") +
  theme(panel.spacing = unit(.01, "in"))
```

Show synaptic input (current) by orientation for simulated V1 neurons on days zero. The resulting dataframe will be used to generate final figures.

```{r, fig.height=2, fig.width=1.4}
I_original = np$dot(y_LGN, w_LGN_to_V1_original)
I_final = np$dot(y_LGN, w_LGN_to_V1)

df_original_I = 
  I_original %>%
  as.data.frame() %>%
  mutate(Theta = all_theta) %>%
  gather(Cell, Rate, -Theta) %>%
  mutate(Day = "0")

df_final_I = 
  I_final %>%
  as.data.frame() %>%
  mutate(Theta = all_theta) %>%
  gather(Cell, Rate, -Theta) %>%
  mutate(Day = "60")

df_I = bind_rows(df_original_I, df_final_I)

# filter out neurons that weren't active on the last day
df_I %<>% 
  group_by(Cell, Theta, Day) %>%
  mutate(Optimal_phase = Rate == max(Rate)) %>%
  filter(Optimal_phase) %>%
  ungroup() 

# show some example neurons
g_curves_I =
  df_I %>%
  filter(Cell %in% paste("V", V1_cells, sep="")) %>%
  mutate(Degrees = (Theta + pi) * (360/(2*pi)) - 90) %>%
  ggplot(aes(Degrees, Rate, group=Day, color=Day)) +
  facet_grid(Cell~.) +
  geom_line(stat="summary", fun.data="mean_se") +
  scale_x_continuous(breaks=c(0, 45, 90, 135, 180)) +
  scale_color_manual(values=c("blue3", "red3")) +
  ylab("Synaptic input")

g_curves_I
```

