
This file is meant to be run after `Orientation Tuning Notebook.Rmd` and `Extract_curves.Rmd` and relies on variable and dependencies described and instantiated there.

Generate firing rate and synaptic input by orientation curves using data extracted in `Extract_curves.Rmd`.

```{r}
# firing rate curves
g_curves = 
  df %>%
  mutate(Day = ifelse(Day=="60", "118", Day)) %>%
  filter(Cell %in% paste("V", V1_cells, sep="")) %>%
  mutate(Degrees = (Theta + pi) * (360/(2*pi)) - 90) %>%
  mutate(Cell = Cell %>% readr::parse_number() %>% as.character()) %>%
  ggplot(aes(Degrees, Rate, group=Day, lty=Day)) +
  facet_grid(Cell~., labeller=label_both) +
  geom_line(stat="summary", fun.data="mean_se") +
  ylab("Firing rate") +
  xlab("Stim. orientation [°]") +
  theme(legend.position="top") +
  guides(lty=guide_legend(keywidth = .5))

# synaptic input curves
g_curves_I =
  df_I %>%
  mutate(Day = ifelse(Day=="60", "Day: 118", "Day: 0")) %>%
  filter(Cell %in% paste("V", V1_cells, sep="")) %>%
  mutate(Degrees = (Theta + pi) * (360/(2*pi)) - 90) %>%
  mutate(Cell = Cell %>% readr::parse_number() %>% as.character()) %>%
  ggplot(aes(Degrees, Rate, group=Day, lty=Day)) +
  facet_grid(Cell~., labeller=label_both) +
  geom_line(stat="summary", fun.data="mean_se") +
  ylab("Synaptic input") +
  xlab("Stim. orientation [°]") +
  theme(legend.title = element_blank())
```

Plot histograms of synaptic strengths across all LGN to V1 synapses on the first and last simulated days.

```{r}
df_strengths =
  bind_rows(
    w_LGN_to_V1_original %>%
      data.frame() %>%
      gather(Cell, Strength) %>%
      filter(Strength > 0) %>%
      mutate(Day = "Original"),
    w_LGN_to_V1 %>%
      data.frame() %>%
      gather(Cell, Strength) %>%
      filter(Strength > 0) %>%
      mutate(Day = "Final")
  ) %>%
  mutate(Day = factor(Day, levels=c("Original", "Final")))

g_strengths = 
  df_strengths %>%
  ggplot(aes(Strength)) +
  facet_grid(Day~.) +
  geom_histogram(bins=100, color="black", fill="black") +
  ylab("Count") +
  xlab("Synaptic strength [au]")

g_strengths
```

